Formalizing Stack Safety as a Security Property

### 2021-07-07 submission3/main.tex
CP

StkTokens \citep{SkorstengaardSTKJFP} continues this work by defining a more efficient
calling convention that makes use of linear capabilities for stack and return
pointers.
%% The convention operates on a single shared stack and requires that
%% protected components avoid compromising their own security by following certain
%% simple rules---like not leaking their private capability seals. More
%% precisely but
Still informally, local state encapsulation (LSE) is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow (WBCF) as only allowing returns from the topmost
frame to the immediately adjacent frame below. Formally, it improves on
\citet{SkorstengaardLocal} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return instructions,
which is proven fully abstract with respect to a more concrete capability
machine.
%% that replaces those pseudo-instructions with their calling convention.
Their proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.
%% ; their model of
%% observations is limited to cotermination.

We improve upon \citet{SkorstengaardSTKJFP} in two ways.

\begin{enumerate}[label=(\roman*)]
\item We capture a desirable security characteristic of dynamically nested
      calls: that {\em every} caller should be protected from its callees.
      By contrast, they statically divide the world into
      trusted functions, which feature additional dynamic "reasonableness"
      constraints on their behavior and for which their properties hold,
      and adversary functions, which may violate those constraints but
      are not guaranteed safety.

      Supporting a dynamic notion of adversary is important for
      functions that cannot be statically proven safe, but whose data
      still need to be protected when they are behaving safely.
      For example, consider a function that obeys the reasonableness
      constraints on most inputs, but can be subverted by malicious inputs
      in such a way that it violates the reasonableness constraints. In
      the StkTokens model, we cannot guarantee to protect this function's
      data from its callees even when it called with non-malicious inputs.
      \apt{too abstract now?}
      \sna{Maybe, but I'm not sure how much being concrete about it serves us.}

\item Their approach to proving safety via fully abstract overlay
  semantics relies on the definition of a ``stack-safe-by-construction''
  abstract machine that is closely tied to the underlying actual
  machine architecture and and enforcement mechanism.
  %the the abstract machine being architecturally similar to
  %      the concrete one (beyond the specific call and return semantics that
  %      distinguish them).
  While the fundamental approach can no doubt be ported to other architectures and enforcement
  schemes, each such port requires a new abstract machine definition whose
  safe-by-construction character must be (informally) verified afresh.
  Our properties are more generic: since we abstract over most of the machine
  state as well as the sealing and return conventions, they apply directly
  to a range of architectures, code features, and enforcement mechanisms.

  Likewise, it would be possible to extend their abstract machine to pass arguments
  on the stack, or to support callee-save registers. It might even viable to
  construct a safe-by-construction abstract coroutine machine, but so far this is
  hypothetical. We implement these scenarios in fact, with simple modifications to
  the domain model of each one.

\end{enumerate}

@@

For criterion (1), we believe our property definitions are at least as intuitive
as theirs.
We use the language of formal security
properties to apply well-established concepts and formal techniques
of non-interference. Thus, we define our properties in terms of
traces of memory and register states, rather than in terms of an
ad-hoc abstract machine.
Which approach is better may be a matter of taste, but we believe that our
alternative has the merit of explicitly relating stack safety to existing
notions of security.

With respect to criterion (4), we believe that our support
for dynamically changing notion of adversaries, as outlined in point (i) above,
captures a better definition of safety than LSE; moreover, our observational
properties capture a richer and more intensional notion of stack safety
that can be applied to a broader range of enforcement mechanisms and code features.
As to criterion (5), we have demonstrated that our approach extends
to coroutines, which is a first step towards multithreading.
Dynamic code generation is more challenging, but StkTokens does not yet
handle this either.
StkTokens is specifically set up to address criterion (3), while this is not
a particular goal of our work. We do agree that criterion (2) is important,
and this remains as future work for us.
Finally, we believe that having a generic definition of stack
safety that abstracts over machine details is also
an important criterion, and here we do better than StkTokens, as described
in point (ii) above.

@@

We have evaluated {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
  Clearing} against our properties, finding that {\em Depth Isolation} does
indeed enforce stepwise stack safety and therefore also observable stack
safety. We also found that neither lazy micro-policy actually does enforce
stack safety, and proposed a fix.

@@

Heap safety, like stack safety, may be framed as a security property
in the form of
noninterference~\citep{DBLP:conf/post/AmorimHP18}\ifaftersubmission\bcp{Should
be spelled ``Azedevo et al.''  Got to figure out how to make bibtex do that
(or ask arthur how!).}\fi. Just
as~\citeauthor{DBLP:conf/post/AmorimHP18} give a rigorous
characterization of the meaning of (heap) memory safety, in this paper,
inspired by their work, we aim to do the same for the stack. They are a
significant inspiration for this work.

@@

We have introduced a characterization against which stack safety enforcement
mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.
Our immediate next step is to implement testing of observational properties,
both to validate them and to explore how to effectively test properties that do not
fail immediately.
From there there are two natural ways to extend the work. We should thoroughly evaluate
other mechanisms, especially Cheri-based capability systems. And we can further demonstrate
the functionality of our definitions by showing that they can be used to prove
the correctness of a programming logic with strengthened rules around calls and returns.

>>>>

StkTokens \citep{SkorstengaardSTKJFP} continues this line of work by
defining a more efficient
calling convention that uses linear capabilities for stack and return
pointers.
%% The convention operates on a single shared stack and requires that
%% protected components avoid compromising their own security by following certain
%% simple rules---like not leaking their private capability seals. More
%% precisely but
Informally, local state encapsulation (LSE) is defined as
restricting accesses to the range of memory allocated to the current stack
frame, and well-bracketed control flow (WBCF) is defined as only allowing
returns from the topmost
frame to the immediately adjacent frame below. Formally, this work improves on
\citet{SkorstengaardLocal} by building those properties into the semantics of a
capability machine with a built-in call stack and call and return
instructions; this machine is proven fully abstract with respect to a more
concrete capability machine.
%% that replaces those pseudo-instructions with their calling convention.
The proof of full abstraction uses a standard notion of components, which
import and export functions through their interfaces.

We improve upon \citet{SkorstengaardSTKJFP} in two significant ways:
\begin{enumerate}[label=(\roman*)]
\item Most importantly, our definition captures the intuition that {\em
  every} caller should be protected from its callees.  By contrast,
\citet{SkorstengaardSTKJFP}
statically divide the world into trusted functions, which feature additional
dynamic ``reasonableness'' constraints on their behavior and for which their
properties hold, and adversary functions, which may violate those
constraints but are not guaranteed safety.  \bcp{Readers may wonder: why
  isn't this reasonable?  after all, if a
  function is adversarial, why do we care whether it gets any safety
  guarantees?  This is an opportunity to (re)explain that functions can be
  wrong without being malicious.  The next paragraph is a step in that
  direction, but I think the emphasis on proof gives the wrong idea.}

\vspace{.7\smallskipamount}

Supporting a dynamic notion of adversary is important for functions that
cannot be statically proven safe, but whose data still needs to be protected
when they are behaving safely.  For example, consider a function that obeys
the reasonableness constraints on most inputs, but can be subverted by
malicious inputs in such a way that it violates the reasonableness
constraints. In the StkTokens model, we cannot guarantee to protect this
function's data from its callees even when it called with non-malicious
inputs.

      % \apt{too abstract now?}
      % \sna{Maybe, but I'm not sure how much being concrete about it serves us.}
      % BCP: reads OK to me

\item \citet{SkorstengaardSTKJFP}'s approach to proving safety via fully
abstract overlay
  semantics relies on the definition of a ``stack-safe-by-construction''
  abstract machine that is closely tied to an underlying enforcement
  mechanism and actual machine architecture.
  %the the abstract machine being architecturally similar to
  %      the concrete one (beyond the specific call and return semantics that
  %      distinguish them).
  While the fundamental approach can no doubt be ported to other architectures and enforcement
  schemes, each such port will require a new abstract machine definition whose
  safety-by-construction property must be checked afresh.
  Our properties are more generic: since we abstract over most of the machine
  state as well as the sealing and return conventions, they apply directly
  to a range of architectures, code features, and enforcement mechanisms.

\vspace{.7\smallskipamount}

  Conversely, it seems possible to extend \citet{SkorstengaardSTKJFP}'s
  abstract machine to pass
  arguments on the stack, or to support callee-save registers. It might even
  viable to construct a safe-by-construction abstract coroutine machine, but
  so far this is hypothetical. We implement these scenarios in fact, with
  simple modifications to the domain model of each one.  \bcp{Don't
    understand the last sentence}

\ifcomments \item \bcp{Should we also mention our testing stuff?} \fi
\end{enumerate}

@@

For criterion (1), we contend that our property definitions are at least as
intuitive as theirs, while being phrased in interestingly different terms.
We use the language of formal security
properties to apply well-established concepts and formal techniques
of non-interference, and we define our properties in terms of
traces of memory and register states, rather than in terms of an
ad-hoc abstract machine.
Which approach is ``better'' is probably a matter of taste, but ours
has at least the merit of explicitly relating stack safety to existing
notions of security.

With respect to criterion (4), we believe that our support
for dynamically changing notion of adversaries, as outlined in point (i) above,
captures a better definition of safety than LSE; moreover, our observational
properties capture an even richer and more intensional notion of stack safety
that can be applied to a broader range of enforcement mechanisms and code features.

As for criterion (5), we have demonstrated that our approach extends
to coroutines, a first step towards multithreading.
Dynamic code generation is more challenging, but StkTokens does not yet
handle this either.
StkTokens is specifically set up to address criterion (3), while this is not
a goal of our work. We do agree that criterion (2) is important;
demonstrating it for our proposal remains future work.

 Finally, we would propose one more criterion: having a generic definition
 of stack safety that abstracts over machine details.  Here we do better
 than StkTokens, as described in point (ii) above.

@@

We evaluated {\em Depth Isolation}, {\em Lazy Tagging}, and {\em Lazy
  Clearing} against our properties, finding that {\em Depth Isolation} does
indeed enforce stepwise stack safety and therefore also observable stack
safety. We also found that neither lazy micro-policy actually enforces
stack safety, and we proposed a fix.

@@

Heap safety, like stack safety, may be framed as a security property using
the language of
noninterference~\citep{DBLP:conf/post/AmorimHP18}\iftrue\bcp{Should be
  spelled ``Azedevo et al.''  We should figure out how to make bibtex do
  that (or ask arthur how!).}\fi.  Their paper was a significant inspiration
for the present work, and we believe that the two safety conditions should
combine smoothly.

@@

We introduced a characterization against which stack safety enforcement
mechanisms can be measured and used it to evaluate tag-based enforcement mechanisms.

\bcp{Out of date?}Our immediate next step is to implement testing of observational properties,
both to validate them and to explore how to effectively test properties that do not
fail immediately.
From there there are two natural ways to extend the work. We should thoroughly evaluate
other mechanisms, especially Cheri-based capability systems. And we can further demonstrate
the functionality of our definitions by showing that they can be used to prove
the correctness of a programming logic with strengthened rules around calls and returns.
